---
title: "Generalized SIR and HawkesN processes"
output: github_document
---

## Introduction

This repo hosts the code or links for reproducing the experiments presented in [1]. Specifically, we summarize the necessary components here:

  - Datasets: we used three large Twitter retweet cacade datasets in our experiments. These datasets can be found at the links: [ActiveYT](https://github.com/computationalmedia/hip-popularity/tree/master/data), [NEWS](https://github.com/computationalmedia/featuredriven-hawkes/tree/master/data) and [SEISMIC](http://snap.stanford.edu/seismic/).
  - Generalized SIR Simulation: code appeared in this repo at `R/sir-simulation.R`.
  - HawkesN Simulation and Fitting: code available in the package [HawkesFit](https://github.com/qykong/HawkesFit).
  - HawkesN Goodness-of-fit Tests: code appeared in this repo at `R/goodness-of-fit.R`.

## Tutorial 1: fitting HawkesN on simulated generalized SIR processes

In this tutorial, we are going to first simulate some stochastic SIR processes with an exponential recovery distribution and then fit them using EXPN (i.e. HawkesN processes with an exponential kernel).

Let's first load all required packages and functions.
```{r, include=FALSE}
source('R/package.R')
source('R/goodness-of-fit.R')
source('R/sir-simulation.R')
set.seed(2)
```

We then simulate some stochastic SIR processes
```{r,message=FALSE}
sim_no <- 50 # simulate 100 processes
params.S <- c(beta = 2, gamma = 0.5, I.0 = 1, N = 100) # chosen SIR parameters 
sims <- lapply(seq(sim_no), function(i) {
  # let's hide the printing from generate.stochastic.sir
  sink(tempfile()) 
  on.exit(sink())
  generate.stochastic.sir(params = params.S, distribution.type = 'EXP')
})

head(sims[[1]])
```
A simulated process is represented by a `data.frame` where each row is an event, i.e. an infection or a recovery event. The columns are event times, suscetible counts $S_t$, infected count $I_t$, recovered counts $R_t$ and cumulative infected counts $C_t$, respectively. For example, the row $6$ represents an recovery event at time $1.5091091$ and the process stopped as there is no infected individual in the population anymore.

Before fitting an EXPN on these simulated processes, we first need to convert SIR processes to the Hawkes-friendly format, i.e. dropping all recovery events
```{r}
sims.H <- lapply(sims, SIR2HAWKES.stochastic.event.series)

head(sims.H[[1]])
```
The converted `data.frame` now only keeps the infection event timestamps.

Let's now fit the EXPN model
```{r}
fit_series(sims.H, model_type = 'EXPN', cores = 10, observation_time = Inf, lower_bound = c(N = 100), upper_bound = c(N = 100))
```
## Reference

If you find the code helpful, please consider citing the paper:

> [1] Kong, Quyu, Marian-Andrei Rizoiu, and Lexing Xie. "Modeling Information Cascades with Self-exciting Processes via Generalized Epidemic Models." arXiv preprint arXiv:1910.05451 (2019).